A Logical Solution of the Linearity Challenge

### 2025-06-19 paper/calculus.tex
++

The calculus of pure sessions that is formalized in our development is closely
related to the multiplicative, additive fragment of CP~\citep{Wadler14}, without
shared channels. Its syntax makes use of an infinite set of \emph{channel
names}, ranged over by $x$, $y$ and $z$, and is shown in \cref{tab:syntax}.
%
A link $\Link\x\y$ denotes the merging of the $x$ and $y$ channels, so that
every message sent on one of the channels is forwarded to the other. This form
is useful for modeling \emph{free outputs}, as we will see later.\Luca{Mettere
esempio!}
%
The $\Close\x$ and $\Wait\x.P$ processes respectively model the output and input
of a termination signal on the channel $x$. The former process has no
continuation, while the latter process continues as $P$ once the session has
been terminated.
%
The $\Fail\x$ process denotes a runtime error concerning the session $x$.
%
The $\Select\x{\InTag_i}\z.P$ and $\Case\x\z{Q_1}{Q_2}$ processes respectively
denote the output and input of a label $\InTag_i$ along with a fresh
continuation $z$ on the channel $x$. Depending on the label $\InTag_i$, the
receiver continues as $Q_i$. In practice it is useful to consider arbitrary
(domain-specific) sets of tags instead of the canonical two $\InTag_1$ and
$\InTag_2$. This generalization has no substantial impact on the formalization,
so we stick to the classical case of ``binary choices'' in line with the usual
presentations of linear logic.
%
The $\Fork\x\y\z{P}{Q}$ and $\Join\x\y\z.R$ processes respectively denote the
output and input of a pair of fresh channels $y$ and $z$ on the channel $x$.
These forms are often used in conjunction with links to model \emph{free
outputs}, but can also be used to model bifurcating sessions whereby a
sequential protocol continues as two parallel sub-protocols.

>>>>

The calculus of pure sessions that is formalized in our development is dubbed
\Calculus and is closely related to the multiplicative, additive fragment of
CP~\citep{Wadler14}, without shared channels. Its syntax makes use of an
infinite set of \emph{channel names}, ranged over by $x$, $y$ and $z$, and is
shown in \cref{tab:syntax}.
%
A link $\Link\x\y$ denotes the merging of the $x$ and $y$ channels, so that
every message sent on one of the channels is forwarded to the other. This form
is useful for modeling \emph{free outputs}, as we will see later.\Luca{Mettere
esempio!}
%
The $\Close\x$ and $\Wait\x.P$ processes respectively model the output and input
of a termination signal on the channel $x$. The former process has no
continuation, while the latter process continues as $P$ once the session has
been terminated.
%
The $\Fail\x$ process denotes a runtime error concerning the session $x$.
%
The $\Select\x{\InTag_i}\z.P$ and $\Case\x\z{Q_1}{Q_2}$ processes respectively
denote the output and input of a label $\InTag_i$ along with a fresh
continuation $z$ on the channel $x$. Depending on the label $\InTag_i$, the
receiver continues as $Q_i$. In practice it is useful to consider arbitrary
(domain-specific) sets of tags instead of the canonical two $\InTag_1$ and
$\InTag_2$. This generalization has no substantial impact on the formalization,
so we stick to the classical case of ``binary choices'' in line with the usual
presentations of linear logic.
%
The $\Fork\x\y\z{P}{Q}$ and $\Join\x\y\z.R$ processes respectively denote the
output and input of a pair of fresh channels $y$ and $z$ on the channel $x$.
These forms are often used in conjunction with links to model \emph{free
outputs}, but can also be used to model bifurcating sessions whereby a
sequential protocol continues as two parallel sub-protocols.
